---
title: "Chapter 4 The tidyverse"
author: "Philwoo, Han"
output:
  html_document:
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: inline
---

Up to now we have been manipulating vectors by reordering and subsetting them through indexing. However, once we start more advanced analyses, the preferred unit for data storage is no the vector but the data frame. In this chapter we learn to work directly with data frames, which greatly facilitate the organization of information. We will be using data frames for the majority of this book. We will focus on a specific data format referred to as *tidy* and on specific collection of packages that are particularly helpful for working with *tidy* data referred to as the *tidyverse*
```{r}
library(tidyverse)
```
We will learn how to implement the tidyverse approach throughout the book, but before delving into the details, in this chapeter we introduce some of the most widely used tidyverse functionality, starting with the **dplyr** package for manipulating data frames and the **purrr** package for working with functions. Note that the tidyverse also includes a graphing package, **ggplot2**, which we introduce later in chapter 7 in the Data Visualization part of the book; the **readr** package discussed in Chapter 5; and many others. In this chapter, we first introduce the concept of *tidy data* and then demonstrate how we use the tidyverse to work with data frames in this format.

## 4.1 Tidy data
We say that a data table is in *tidy* format if each row represents one observation and columns represent the different variables available for each of these observations. The murders dataset is an example of a tidy data frame
```{r}
library(dslabs)
data(murders)
```
Each row represent a state with each of the five columns providing a different variable related to these states: name, abbreviation, region, population, and total murders.
To see how the same information can be provided in different formats, consider the following example:
This tidy dataset provides fertility rates for two countries across the years. This is a tidy dataset because each row presents one observation with the three variables being country, year, and fertility rate. However, this dataset originally came in another format and was reshped for the **dslabs** package. Originally, the data was in the following format:
The same information is provided, but there are two improtant differences in the format: 1) each row includes several observations and 2) one of the variables, year, is stored in the header. For the tidyverse packages to be optimally used, date need to be reshped into tidy format, which you will learn to do in the Data Wrangling part of the book. Until then, we will use example datasets that are already in tidy format.
Although not immediately obvious, as you go through the book you will start to appreciate the advantages of working in a frame work in which functions use tidy formats for both inputs and outputs. You will see how this permits the data analyst to focus on more important aspects of the analysis rather than the format of the data.

##4.2 Exercises

1. Examine the built-in dataset co2. Which of the following is true:
```{r}
data(co2)
str(co2)
```
d. co2 is not tidy: to be tidy we would have to wrangle it to have three columns (year, month and value), then each co2 observation would have a row.
2. Examine the built-in dataset ChickWeight. Which of the following is true:
```{r}
data(ChickWeight)
str(ChickWeight)
```
d. ChickWeight is tidy: it is stored in a data frame
3. Examin the built-in dataset BOD. Which of the following is true:
```{r}
data(BOD)
```
c. BOD is tidy: each row is an observation with two values(time and demand)
4. Whcih of the following built-in datasets is tidy(you can pick more than one):
```{r}
data(BJsales)
data(EuStockMarkets)
data("DNase")
data(Formaldehyde)
data(Orange)
data(UCBAdmissions)
```
b. EuStockMarkets
c. DNase
d. Formaldehyde
e. Orange
f. UCBAdmissions

## 4.3 Manipulating data frames

The **dplyr** package from the **tidyverse** introduces functions that perform some of the most common operations when working with data frames and uses names for these functions that are relatively easy to remember. For instance, to change the data table by adding a new column, we use mutate. To filter the data table to a subset of rows, we use filter. Finally, to subset the data by selecting specific columns, we use select.

### 4.3.1 Adding a Column with mutate

We want all the necessary information for our analysis to be included in the data table. So the first task is to add the murder rates to our murders data frame. The function mutate takes the data frame as a first argument and the name and values as a second argument using the convention name - values. So, to add murder rates, we use:
```{r}
library(dslabs)
data("murders")
murders <- mutate(murders, rate = total / population *100000)
```
Notice that here we used total and population inside the function, which are objects that are **not** defined in our workspace. But why don't we get an error?
This is one of **dplyr**'s main features. Functions in this package, such as mutate, know to look for variables in the data frame provided in the first argument. In the call to mutate above, total will have the values in murders$total. This approach makes the code much more readable.
We can see that the new column is added:
```{r}
head(murders)
```
Although we have overwritten the original murders object, this does not change the object that loaded with data(murders). If we load the murders data again, the original will overwrite our mutated version.

### 4.3.2 Subsetting with filter

Now suppose that we want to filter the data table to only show the entries for which the murder rate is lower than 0.71. To do this we use the filter function, which takes the data table as the first argument and then the conditional statement as the second. Like mutate, we can us the unquoted variable names from murders inside the function and it will know we mean the columns and not objects in the workspace.
```{r}
filter(murders, rate <= 0.71)
```

```{r}
murders
```

