---
title: "Chapter 2"
date: 2021/09/07"
name: "Philwoo, Han"
output: html_notebook
---
# Chapter 2 R basics

In this book, we will be using the R software environment for all our analysis. You will learn R and data analysis techniques simultaneously. To follow along you will therefore need access to R. We also recommend the use of an *integrated development environment*(IDE), such as Rstudio, to save your work. Note taht it is common for a course or workshop to offer access to an R environment and an IDE through your web browser, as done by RStudio cloud. If you have access to such a resource you don't need to install R and RStudio. However, if you intend on becoming an advanced data analyst, we highly recommend installing these tools on your computer. Both R and RStudio are free and available online.

## 2.1 Case study: US Gun Murders

Imagine you live in Europe and are offered a job in a US company with many locations across all states. It is a great job, but news with headlines such as **US Gun Homicide Rate Higher Than Other Developed Countries** have you worried. 
Charts like this may concern you even more:
Or even worse, this version from [everytown.org](http://www.everytown.org):
But then you remeber that the US is a large and diverse country with 50 very different states as well as the District of Columbia. California, for example, has a larger population than Canada, and 20 US states have populations larger than that of Norway. In some respects, the variability across states in the US is akin to the variability across countries in Europe. Furthermore, although not included in the charts above, the murder rates in Lithuania, Ukraine, and Russia are higher than 4 per 100,000. So perhaps the new reports that worried you are too superficial. You have options of where to live and want to determine the safety of each particular state. We will gain some insights by examining data related to gun homicides in the US during 2010 Using R.
Before we get started with our example, we need to cover logistics as well as some of the very basic building blocks that are required to gain more advanced R skills. Be aware that the usefulness of some of these building blocks may not be immediately obvious, but later in the book you will appreciate having mastered these skills.

## 2.2 The very basics

Before we get started with the motivating dataset, we need to cover the very basics of R.

### 2.2.1 Objects

Suppose a high school student asks us for help solving several quadratic equations of the form $ax^2+bx+c = 0$. The quadratic formula gives us the solutions: $$\frac{-b - \sqrt{b^2 - 4ac}}{2a}\,\, \mbox{ and } \frac{-b + \sqrt{b^2 - 4ac}}{2a}$$ which of course change depending on the values of *a,b* and *c*. One advantage of programming languages is that we can define variables and write expressions with these variables, similar to how we do so in math, but obtain a numeric solution. We will write out general code for the quadratic equation below, but if we are asked to solve $x^2 + x -1 = 0$, than we define:
```{r}
a <- 1
b <- 1
c <- -1
```
which store the values for later use, We use <- to assign values to the variables.
We can also assign values using = instead of <-, but we recommend against using - to avoid confusion.
Copy and paste the code above into your console to define the three variables. Note that R does not print anything when we make this assignment. This means the objects were defined successfully. Had you made a mistake, you would have received an error message.
To see the value stored in a variable, we simply ask R to evaluate a and it shows the stored value:
```{r}
a
```
A more explicit way to ask R to show us the value stored in a is using print like this
```{r}
print(a)
```
We us the term *object* to describe stuff that is stored in R. Variables are examples, but objects can also be more complicated entities such as functions, which are described later.

### 2.2.2 The workspace

As we define objects in the console, we are actually changing the **workspace**. You can see all the variables saved in your workspace by typing:
```{r}
ls()
```
In RStudio, the *Environment* tab shows the values:
We should see a, b and c. If you try to recover the value of a variable that is not in your workspace, you receive an error. For example, if you type x you will receive the following message: Error: object 'x' not found.
Now since these values are saved in variables, to obtain a solution to our equation, we use the quadratic formula:
```{r}
(-b + sqrt(b^2 - 4*a*c)) / (2*a)
(-b - sqrt(b^2 - 4*a*c)) / (2*a)
```

### Functions

Once you define variables, the data analysis process can usually be described as a series of *functions* applied to the data. R includes several predefined functions and most of the analysis pipelines we construct make extensive us of these
We already use the install.packages, library, and ls functions. We also used the function sqrt to solve the quadrtic equation above. There are many more prebuilt functions and even more can be added through packages. These functions do not appear in the worspace because you did not define them, but they are available for immediate use.
In general, we need to use parentheses to evaluate a function. If you type ls, the functions is not evaluated and instead R shows you the code that defines the function. If you type ls() the function is evaluated and as seen above, we see objects in the workspace.
Unlike ls, most functions require one or more *arguments*. Below is an example of how we assign an object to the argument of the function log. Remember that we earlier defined a to be 1:
```{r}
log(8)
log(a)
```
You can find out what the function expects and what it does by reviewing the very useful manuals included in r. you can get help by using the help function like this:
```{r}
help("log")
```
For most functions, we can also use this shorthand:
```{r}
?log
```
This help page will show you what arguments the function is expecting. For example, log needs x and base to run However, some arguments are required and others ar optional. You can determine which arguments are optional by noting in the help document that a default value is assinged with =. Defining these is optional. For example, the base of the function log defaults to base = exp(1) making log the natural log by default.
If you want a quick look at the qrguments without opening the help system, you cna type:
```{r}
args(log)
```
You cna change the default values by simply assigning another object:
```{r}
log(8, base = 2)
```
Note that we have no been specifying the argument x as such:
```{r}
log(x = 8, base = 2)
```
The above code works, but we can save ourselves some typing: if no argument name is used, R assumes you are entering arguments in the order shown in the help file or by args. So by not using the names, it assumes the arguments are x followed by base :
```{r}
log(8,2)
```
If using the arguments' names, then we can include them in whatever order we want:
```{r}
log(base = 2, x = 8)
```
To specify arguments, we must use =, and cannot use <-. 
There are some exceptions to the rule that functions need the parentheses to be evaluated. Among these, the most commonly used are the arithmetic and relational operators. For example:
```{r}
2^3
```
You can see the arithmetic operators by typing:
```{r}
help("+")
```
or
```{r}
?"+"
```
and the relational operators by typing
```{r}
help(">")
```
or
```{r}
?">"
```

### 2.2.4 Other prebuilt objects

There are several datasets that are included for users to practice and test our functions. You can see all the available datasets by typing:
```{r}
data()
```
This shows you the object name for these datasets. These datasets are objects that can be used by simply typing the name. For example, if you type
```{r}
co2
```
R will show you Mauna Loa atmospheric CO2 concentration data.
Other prebuilt objects are mathematical quantities, such as the constant \pi and \inf:
```{r}
pi
Inf+1
```

### 2.2.5 Variable names

We have used the letters a,b, and c as variable names, but variable names can be almost anything. Some basic rules in R are the variable names have to start with a letter, cna't contain spaces, and should not be variables that are predefined in R. For example, don't name on of your variables install.packages by typing something like install.packages <- 2.
A nice convention to follow is to use meaningful words that describe what is stored, use only lower case, and use underscores as a substitute for spaces. For the quadratic equations, we could use something like this:
```{r}
solutions_1 <- (-b + sqrt(b^2 - 4*a*c)) / (2*a)
soultions_2 <- (-b - sqrt(b^2 - 4*a*c)) / (2*a)
```
For more advice, we highly recommend studying Hadley Wickham's style guide.

### 2.2.6 Saving your workspace
```{r}

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
